\documentclass[12pt]{letter}
\usepackage{geometry}
\geometry{letterpaper, margin=1in}
\usepackage{setspace}
\usepackage{parskip}

\begin{document}

\begin{letter}{
    IC System, Inc. \\
    Correspondence Department \\
    P.O. Box 64319 \\
    St. Paul, MN 55164-0319
}

\date{\today}
\signature{Fhaheem Tadamarry}

\opening{Dear IC System Team,}

My name is Fhaheem Tadamarry, and I am writing in regard to a Verizon Wireless collection account that was assigned to IC System (Account Number: 381937729-001-09). I received a letter from your office regarding this debt and paid the full amount of \$82.13 on the same day—March 25, 2025 (Confirmation Number: 22254700).

First, I want to thank you for your timely handling of the matter. I would now like to respectfully request a goodwill deletion of this account from my credit report.

Unfortunately, due to prior experiences with fraudulent debt collection scams, I initially believed the Verizon notice might have been another scam attempt. Once I confirmed the debt was legitimate, I took immediate responsibility and paid it in full the same day. This situation does not reflect my usual financial behavior.

I have maintained a 100\% on-time payment history, and this account is the only negative item on my otherwise clean credit record. Since it has now been paid and resolved, I am asking—solely as a gesture of goodwill—that you consider removing the record from my credit report with the three major bureaus.

I understand that IC System is not obligated to do this. However, I would be deeply grateful for your assistance in helping me restore my credit standing.

Please let me know if there is any additional documentation or information I can provide.

\closing{Sincerely,}

\ps{Enclosure: Payment Confirmation (Confirmation Number: 22254700)}

\end{letter}
\end{document}